---
title: "Final paper KH boxplots"
author: "Katherine"
date: '2022-04-24'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r}
library("tidyverse")
library("ggstatsplot")
library("ggplot2")
library("ggpubr")
library("corrgram")
library("ellipse")
library("mctest")
library("car")
library("PMCMRplus")
library("dplyr")
library("broom")
```

#Reading in data
```{r}
data <- read.csv("Cunning_ED50.csv")

summary(data)
print(data$estimate_ed50)
```
#Inital look at data-boxplot of thermotolerance vs genotype
```{r}
#boxplot(estimate_ed50~geno, data= data)
```
#Models
```{r}
nurs_loc <- tribble(
  ~nursery, ~nurs_lat, ~nurs_lon, 
  "MML", 24.56, -81.40,
  "NSU", 26.12, -80.09,
  "UM", 25.67, -80.09,
  "CRF", 24.98, -80.43,
  "RR", 24.98, -80.43,
  "FWC", 24.66, -81.02
)

subset2 <- data %>% 
  left_join(nurs_loc) %>% 
  group_by(nursery) %>% 
  arrange(-estimate_ed50)  %>%
  mutate(order_ed50 = 1:n()) %>% 
  #mutate(Ranking = case_when(order_ed50 <= 20 ~ "CRF Top 20", #case_when is if statement
                            #TRUE ~ "Others"))) %>%  #then that, Ranking is new column
  filter(order_ed50 <= 20 ) %>% 
  mutate(nursery = factor(nursery, levels = c("MML", "FWC", "RR", "CRF", "UM", "NSU")))

# Test for equal variance across nursery latitude
car::leveneTest(estimate_ed50 ~ as.factor(nurs_lat), data = subset2) %>%
  knitr::kable(caption = "Levene's test for homogeneity of variance")
#failing, each nursery does not have equal variance of Ed50

## Unequal variance among nurseries -- use Welch ANOVA
welch <- oneway.test(estimate_ed50 ~ as.factor(nurs_lat), data = subset2)
welch
tidy(welch) %>% knitr::kable(caption = "Welch's anova for unequal variance among nurseries")
#p-value = 2.2e-16

#LM and ANOVA for top performers looking at source latitude
mod <- lm(estimate_ed50 ~ source_lat, data = subset2)
anova(mod) %>% knitr::kable()
mod #significant pvalue = 0.04

#LM and ANOVA for top performers looking at nursery latitude
mod2 <- lm(estimate_ed50 ~ nurs_lat, data = subset2)
anova(mod) %>% knitr::kable()
mod2 #not significant pvalue = 0.12

test <- subset2 %>% 
  group_by(nursery) %>% 
  do(mod = tidy(lm(estimate_ed50 ~ source_lat, data = .))) %>% 
  tidyr::unnest(mod) %>% 
  mutate(p.est = round(p.value,3))

subset2 %>% 
  ggplot(aes(x = source_lat, y = estimate_ed50)) +
  geom_point(aes(color = nursery)) +
  geom_smooth(method = 'lm', aes(color = nursery))# +
  facet_wrap(~nursery, scales = 'free')

#within nursery latitude, source latitude is significant
#regardless of nl, ed50 and lat don't correlate
```
#Top performers thermotolerance
##plot of thermotolerance vs. top 20 genotype
```{r}
#plot all 5 nurseries ed50 vs nursery latitude
nurs_loc <- tribble(
  ~nursery, ~nurs_lat, ~nurs_lon, 
  "MML", 24.56, -81.40,
  "NSU", 26.12, -80.09,
  "UM", 25.67, -80.09,
  "CRF", 24.98, -80.43,
  "RR", 24.98, -80.43,
  "FWC", 24.66, -81.02
)

subset2 <- data %>% 
  left_join(nurs_loc) %>% 
  group_by(nursery) %>% 
  arrange(-estimate_ed50)  %>%
  mutate(order_ed50 = 1:n()) %>% 
  #mutate(Ranking = case_when(order_ed50 <= 20 ~ "CRF Top 20", #case_when is if statement
                            #TRUE ~ "Others"))) %>%  #then that, Ranking is new column
  filter(order_ed50 <= 20 ) %>% 
  mutate(nursery = factor(nursery, levels = c("MML", "FWC", "RR", "CRF", "UM", "NSU")))


subset2 %>%
  filter(nursery %in% c("CRF", "RR","MML", "NSU", "FWC", "UM")) %>%
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Nursery Thermotolerance Vary by Nursery Latitude?",
        subtitle = "Top 20 Performers",
         x = "Nursery ED50", y= "Nursery Latitude")
ggplot
```

##Boxplot of thermotolerance vs. top 20 genotype,genetic diversity
```{r}
data %>% 
  left_join(nurs_loc) %>% 
  ggplot(aes(x = estimate_ed50, y = nurs_lat)) +
  geom_boxplot(aes(color = nursery))+
  scale_color_brewer(palette = "Dark2")+
  labs(title = "Does Nursery Thermotolerance Vary by Nursery Latitude?",
         x = "Nursery ED50", y= "Nursery Latitude")
```

##Source latitude
##Plot of thermotolerance by source latitude for CRF and RR
```{r}
subdata <- data %>% 
  arrange(-estimate_ed50) %>% #arrange descending order
  group_by(nursery) %>% 
  mutate(order_ed50 = 1:n()) %>% #made new column for order of ed50 for 1 until end
  mutate(Ranking = case_when(order_ed50 <= 20 ~ "Top 20", #case_when is if statement
                             TRUE ~ "Others")) %>% #then that, Ranking is new column 
  filter(Ranking == "Top 20")

#install.packages("ggpmisc")
#Plot of thermotolerance by source latitude for CRF and RR
subdata %>%   
  filter(nursery %in% c("CRF", "RR")) %>% #filtering for only CRF and RR nurseries
  ggplot(aes(y = source_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Source Latitude for CRF and RR?",
       subtitle = "Top 20 Performers",
         x = "ED50", y= "Source Latitude")
  # ggpmisc::stat_fit_glance(method = 'lm', geom="text", 
  #                          aes(label = paste("R-Squared = ", ..rr.label..)),
  #                              label.x.npc = 37, label.y.npc = 24.25)

#Boxplot is better
subdata %>% 
  filter(nursery %in% c("CRF", "RR")) %>% 
  ggplot(aes(y = source_lat, x = nursery)) +
  geom_boxplot(aes(fill = nursery)) + #using fill to fill the 2 inner quartiles of plot
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Latitude for CRF and RR?",
         x = "Nursery", y= "Source Latitude",
       theme(legend.position = "none"))
  #geom_jitter(aes(fill = Ranking), width = 0.3, alpha = 0.8)

##no real trend which makes sense because CRF and RR are almost at the same site
```

##Plot of thermotolerance by source latitude for NSU and MML
```{r}
subdata4 <- data %>% 
  arrange(-estimate_ed50) %>% #arrange descending order
  group_by(nursery) %>% 
  mutate(order_ed50 = 1:n()) %>% #made new column for order of ed50 for 1 until end
  mutate(Ranking = case_when(order_ed50 <= 20 ~ "Top 20", #case_when is if statement
                             TRUE ~ "Others")) #then that, Ranking is new column

#install.packages("ggpmisc")
#Plot of thermotolerance by source latitude
subdata4 %>%   
  filter(Ranking == "Top 20") %>% 
  filter(nursery %in% c("NSU", "MML")) %>% #filtering for only CRF and RR nurseries
  ggplot(aes(y = source_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Source Latitude for NSU and MML?",
        subtitle = "Top 20 Performers",
         x = "ED50", y= "Source Latitude")
  # ggpmisc::stat_fit_glance(method = 'lm', geom="text", 
  #                          aes(label = paste("R-Squared = ", ..rr.label..)),
  #                              label.x.npc = 37, label.y.npc = 24.25)

# #Boxplot is better
# b <- subdata4 %>% 
#   filter(nursery %in% c("NSU", "MML")) %>% 
#   ggplot(aes(y = source_lat, x = estimate_ed50)) +
#   geom_boxplot(aes(fill = nursery)) + #using fill to fill the 2 inner quartiles of plot
#   scale_color_brewer(palette = "Dark2") +
#   labs(title = "Does Thermotolerance Vary by Latitude for NSU and MML?",
#          x = "ed50", y= "Source Latitude",
#        theme(legend.position = "none"))
#   #geom_jitter(aes(fill = Ranking), width = 0.3, alpha = 0.8)
# print(a+b)
```

#Nursery latitude
##Does thermotolerance vary by nursery latitude for NSU and MML?
```{r}
subset2 <- data %>% 
  left_join(nurs_loc) %>% 
  group_by(nursery) %>% 
  arrange(-estimate_ed50) %>% 
  mutate(order_ed50 = 1:n()) %>% 
  #mutate(Ranking = case_when(order_ed50 <= 20 ~ "CRF Top 20", #case_when is if statement
                            #TRUE ~ "Others"))) %>%  #then that, Ranking is new column
  filter(order_ed50 <= 20 ) %>% 
  mutate(nursery = factor(nursery, levels = c("MML", "FWC", "RR", "CRF", "UM", "NSU")))

#install.packages("ggpmisc")
#Plot of thermotolerance by nursery latitude for NSU and MML
subset2 %>%   
  filter(nursery %in% c("NSU", "MML")) %>% #filtering for only CRF and RR nurseries
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Nursery Latitude for NSU and MML?",
        subtitle = "Top 20 Performers",
         x = "ED50", y= "Nursery Latitude")
  # ggpmisc::stat_fit_glance(method = 'lm', geom="text", 
  #                          aes(label = paste("R-Squared = ", ..rr.label..)),
  #                              label.x.npc = 37, label.y.npc = 24.25)

#Boxplot is better
subset2 %>% 
  filter(nursery %in% c("NSU", "MML")) %>% 
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_boxplot(aes(fill = nursery)) + #using fill to fill the 2 inner quartiles of plot
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Nursery Latitude for NSU and MML?",
         x = "ED50", y= "Nursery Latitude",
       theme(legend.position = "none"))
  #geom_jitter(aes(fill = Ranking), width = 0.3, alpha = 0.8)

```

##Does thermotolerance vary by nursery latitude for CRF and RR?
```{r}
subset2 %>%   
  filter(nursery %in% c("CRF", "RR")) %>% #filtering for only CRF and RR nurseries
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Nursery Latitude for CRF and RR?",
        subtitle = "Top 20 Performers",
         x = "ED50", y= "Nursery Latitude")

subset2 %>% 
  filter(nursery %in% c("CRF", "RR")) %>% 
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_boxplot(aes(fill = nursery)) + #using fill to fill the 2 inner quartiles of plot
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Nursery Latitude for CRF and RR?",
         x = "ED50", y= "Nursery Latitude",
       theme(legend.position = "none"))


subset2 %>%   
  filter(nursery %in% c("FWC", "RR")) %>% #filtering for only CRF and RR nurseries
  ggplot(aes(y = nurs_lat, x = estimate_ed50)) +
  geom_point(aes(color = nursery)) + #making ed50 values into Ranking, which will be different colors
  geom_smooth(method = 'lm', aes(color = nursery), se = FALSE)+ #seeing if there is a trend using lm
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Thermotolerance Vary by Nursery Latitude for FWC and RR?",
        subtitle = "Top 20 Performers",
         x = "ED50", y= "Nursery Latitude")
```

#Extra
```{r}
subdata3 <- data %>% 
  arrange(-estimate_ed50) %>% #arrange descending order
  group_by(nursery) %>% 
  mutate(order_ed50 = 1:n()) %>% #made new column for order of ed50 for 1 until end
  mutate(Ranking = case_when(order_ed50 <= 20 ~ "Top 20", #case_when is if statement
                             TRUE ~ "Others")) #then that, Ranking is new column
#Plot of thermotolerance by source latitude for all nurseries
subdata3 %>% 
  ggplot(aes(y = estimate_ed50, x = source_lat)) +
  #geom_boxplot(aes(fill = Ranking)) + #using fill to fill the 2 inner quartiles of plot
  geom_point(aes(color = nursery)) +
  geom_smooth(method = 'lm') +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Does Nursery Thermotolerance Vary by Source Latitude?",
       subtitle = "",
         y = "ED50", x= "Source Latitude")

#Plot of thermotolerance by source latitude for all nurseries minus UM and MML
subdata3 %>% 
  filter(!nursery %in% c("UM", "MML")) %>%
  left_join(nurs_loc) %>% 
  ggplot(aes(y = estimate_ed50, x = source_lat)) +
  #geom_boxplot(aes(fill = Ranking)) + #using fill to fill the 2 inner quartiles of plot
  geom_point(aes(color = nursery)) +
  geom_smooth(method = 'lm') +
  scale_color_brewer(palette = "Dark2") 


#FWC ed50 isn't as good, not due to source locations or nurs lat
subdata3 %>% 
  filter(nursery %in% c("FWC", "NSU"),
         Ranking == "Top 20") %>% 
  ggplot(aes(y = source_lat, x = nursery)) +
  geom_point(aes(fill = Ranking)) + #using fill to fill the 2 inner quartiles of plot
  geom_point(aes(color = nursery)) +
  geom_smooth(method = 'lm') +
  scale_color_brewer(palette = "Dark2")

#
test2 <- subdata3 %>% 
  filter(nursery %in% c("FWC", "NSU")) %>% 
  group_by(nursery,Ranking) %>% 
  do(mod = tidy(lm(estimate_ed50 ~ source_lat, data = .))) %>% 
  tidyr::unnest(mod) %>% 
  mutate(p.est = round(p.value,3))
``` 




